---
title: "Data import"
---


## Solutions


```{r}
library(tidyverse)
```

## Exercise 8.2.4

1.What function would you use to read a file where fields were separated with "\|"?

read_delim

2.Apart from `file`, `skip`, and `comment`, what other arguments do [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) and [`read_tsv()`](https://readr.tidyverse.org/reference/read_delim.html) have in common?

they have almost all of each others arguments

3.What are the most important arguments to [`read_fwf()`](https://readr.tidyverse.org/reference/read_fwf.html)?

i don't know i've never used it

4.Sometimes strings in a CSV file contain commas. To prevent them from causing problems, they need to be surrounded by a quoting character, like `"` or `'`. By default, [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) assumes that the quoting character will be `"`. To read the following text into a data frame, what argument to [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) do you need to specify?

colnames = false

5.Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\1")
read_csv("a,b\n1,2\na,b")
read_csv2("a;b\n1;3")
```

6.Practice referring to non-syntactic names in the following data frame by:

a.  Extracting the variable called `1`.

b.  Plotting a scatterplot of `1` vs. `2`.

c.  Creating a new column called `3`, which is `2` divided by `1`.

d.  Renaming the columns to `one`, `two`, and `three`.

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

```{r}
annoying %>% rename(one = `1`,two = `2`) %>%  
  mutate( three = two/one) %>% ggplot(aes(x = one,y = two)) +
  geom_point()
```
