---
title: "Exploratory data analysis"
---

## Solutions

```{r}
library(tidyverse)
library(ggbeeswarm)
library(lvplot)
```

## Exercise 12.3.3

1.  look up help ?diamonds

```{r}
ggplot(diamonds, aes(x = x , y = y)) + geom_point()
```

`x` is the length, `y` is the width, `z` is the depth

2.  

```{r}
ggplot(diamonds, aes(x = price)) + geom_histogram(binwidth = 2000)
```

3.  

```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = .1)
```

4.  It looks like when you use the `xlim` it zooms out but when you use the `ylim` it zooms in.

```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = .1) +
  coord_cartesian(xlim = c(0,50))

ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = .1) + 
  coord_cartesian(ylim = c(0,50))
```

## Exercise 12.4.1

1.  The missing values are removed with a warning.

2.  For mean it determines whether na's should be removed or not and for sum should na values be removed and certain other values.

## Exercise 12.5.1.1

1.  

```{r}
canceled_flights <- nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + (sched_min / 60)
  ) 

ggplot(canceled_flights,aes(x = cancelled)) +
  geom_bar()
```

2.I would say maybe clarity or just straight up price,because the lower quality diamonds correspond to people buying them who don't really know about diamonds or don't care about the quality,theres alot more lower quality diamonds than higher quality.

?diamonds

3.  It switching the x and y its alot faster.

```{r}
ggplot(mpg, aes(x = hwy, y = fct_reorder(class, hwy, median))) +
  geom_boxplot() + coord_flip()
```

4.One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of "outlying values". One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using `geom_lv()` to display the distribution of price vs. cut. What do you learn? How do you interpret the plots?



```{r}
ggplot(diamonds,aes(x = cut,y = price)) +
  geom_lv()
```

5. geom_violin shows a basic understanding of the amount of diamonds while geom_histogram shows the further outliers better and where they are and geom_freqpoly shows the count better

```{r}
ggplot(diamonds,aes(x = price, y = clarity)) +
  geom_violin()
```

```{r}
ggplot(diamonds,aes(x = price)) +
  geom_histogram() +
  facet_wrap(~clarity, ncol = 1, scales = "free_y")
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`
```

```{r}
ggplot(diamonds,aes(x = price, color = clarity)) +
  geom_freqpoly()
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```

6. `geom_beeswarm()` makes it so you have more control over overplotting datasets, while geom_jitter to me seems like just a default that you can use on the go but if you do want more customization ggbeeswarm is better for handling overplotting.

## Exercise 12.5.2.1

1.

```{r}
ggplot(diamonds, aes(x = cut, y = color)) +
  geom_count()
```

2.

```{r}
count(diamonds,color,cut)
 
 ggplot(diamonds,aes( x = color, fill = cut)) +
  geom_bar(position = "fill")
```

3.

```{r}
nycflights13::flights |> 
  summarise(dest,month,year) |> 
  group_by(dest,month)
```

## Exercise 12.5.3.1

```{r}
smaller <- diamonds |> 
  filter(carat < 3)
```

1. If you want to use cut_width know you r datas values,if you use cut_number know your sample size.

```{r}
# visualize price binning by carat, cut_width()
ggplot(smaller, aes(x = price, y = ..density..,)) +
  geom_freqpoly(aes(color = cut_width(carat, 0.5)))

# visualize price binning by carat, cut_number(), 10 bins
ggplot(smaller, aes(x = price, y = ..density..,)) +
  geom_freqpoly(aes(color = cut_number(carat, 10)))
```

2.Visualize the distribution of `carat`, partitioned by `price`.

```{r}
ggplot(diamonds, aes(x = price, y = carat)) + 
  geom_boxplot()
```

3. Its not what i expect even the large diamonds are the same price as smaller diamonds.

```{r}
ggplot(diamonds,aes(x = carat, y = price)) +
  geom_point()
```

4.

```{r}
ggplot(diamonds,aes(x = price, color = cut, fill = cut)) +
  geom_freqpoly()
```

```{r}
ggplot(diamonds,aes(x = carat, color = cut, fill = cut)) +
  geom_freqpoly()
```

5. This scatterplot shows outliers and binned plots don't.

```{r}
diamonds |> 
  filter(x >= 4) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
```

```{r}
diamonds |> 
  filter(x >= 4) |> 
ggplot(aes(x = x)) +
  geom_histogram()
```

6. If you want to control the width you use cut_width, if you want to choose the number of bins use cut_number.

```{r}
ggplot(smaller, aes(x = carat, y = price)) + 
  geom_boxplot(aes(group = cut_number(carat, 20)))

smaller |> 
  mutate(carat_group = cut_number(carat,20)) |>
  count(carat_group)
```

```{r}
ggplot(smaller, aes(x = carat, y = price)) + 
  geom_boxplot(aes(group = cut_width(carat, .1)))

smaller |>
  mutate(carat_group = cut_width(carat,.1)) |> 
  count(carat_group)
```
