---
title: "chapter-13"
editor: visual
---

## Solutions

## Prerequisites

```{r}
library(tidyverse)
library(nycflights13)
```

## Exercises 13.2.4

1.Near basically tells you if a number is near another number, so for example you could ask if 1.99999 is near 2 and it'll say true or false.

```{r}
near(sqrt(2)^2,2)
```

2\. Dep_time and dep_delay has the same amount of NA's because if there wasn't a flight then there wouldn't be a dep_delay.

```{r}
colSums(is.na(flights))
```

## Exercises 13.3.4

1\.

```{r}
flights |>
  mutate(arr_delay1 = is.na(arr_delay))
```

```{r}
flights |>
  mutate(arr_delay1 = is.na(arr_delay)) |>
  select(arr_delay1, arr_time, sched_arr_time)
```

2.8,255 NA's in dep_time, these rows represent cancelled flights.

```{r}
colSums(is.na(flights))
```

3.There is no pattern or connection between the two.

```{r}


# Filter out flights with missing departure times
flights_filtered <- filter(flights, !is.na(dep_time))

# Calculate number of cancelled flights per day
cancelled_flights_per_day <- flights_filtered %>%
  group_by(year, month, day) %>%
  summarise(cancelled_flights = sum(is.na(dep_delay)))

# Calculate proportion of cancelled flights and average delay of non-cancelled flights per day
proportion_delay_per_day <- flights_filtered %>%
  group_by(year, month, day) %>%
  summarise(proportion_cancelled = sum(is.na(dep_delay))/n(),
            avg_delay = mean(dep_delay[!is.na(dep_delay)]))


```

## Exercises 13.4.4

1.It will give you the total amount of NA's in x, and mean would give you the average number of NA's.

2.When prod() is applied to a logical vector, it coerces it to numeric and then returns the product of all the values present in the vector. [If any value is missing or NaN, the result is NaN](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/prod).

The **`min()`** function returns the minimum value of its arguments. When applied to a logical vector, it coerces it to numeric and then returns the minimum value of all the values present in the vector. [If any value is missing or NaN, the result is NaN](https://stackoverflow.com/questions/2190756/how-to-count-true-values-in-a-logical-vector).

Both min and prod are similar to as.numeric.

## Exercises 13.5.4

1\.

```{r}
if_else(x %% 2 == 0, "even", "odd")
```

## 

2\.

```{r}
if_else(y %in% "Saturday", "weekend", "weekday")
```

3\.

```{r}
if_else(abs(x) >= 0, "yes", "no")
```

4\.

```{r}
flights |>
  mutate(holiday = case_when(
    month == 1 & day == 1 ~ "New Years Day",
    month == 7 & day == 4 ~ "4th of July",
    month == 12 & day == 25 ~ "Christmas"
  )) |>
  mutate(yesnoholidays = case_when(
    holiday == "New Years Day" ~ "TRUE",
    holiday == "4th of July" ~ "TRUE",
    holiday == "Christmas" ~ "TRUE"
  ))
```
