---
title: "Communication"
---

```{r}
#| results: "asis"
#| echo: false

source("_common.R")
```

## Prerequisites

```{r}
library(scales)
library(tidyverse)
library(lubridate)
library(ggrepel)
library(patchwork)
library(ggthemes)
```

## 12.2.1 Exercises {.unnumbered}

1.  

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs( title = "insert title here",
        subtitle = "insert subtitle here",
        caption = "insert source here",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    color = "Car type"
  )
```

2.  

```{r}
mpg |> ggplot(aes(x = cty, y = hwy, color = drv, shape = drv)) +
  geom_point() +
  labs(x = "City MPG",
        y = "Highway MPG",
        shape = "Type of drive train",
       color = "Type of drive train"
       )
```

## 12.3.1 Exercises  {.unnumbered}

1.  

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
 annotate(geom = "text",
          x = 7, y = 45, label = "help",
          color = "red")
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
 annotate(geom = "text",
          x = 1, y = 45, label = "help",
          color = "red")
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
 annotate(geom = "text",
          x = 1, y = 1, label = "help",
          color = "red")
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
 annotate(geom = "text",
          x = 7, y = 1, label = "help",
          color = "red")
```

2.  

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  annotate(
    geom = "label", x = 3.5, y = 38,
    label = "Increasing engine size is \nrelated to decreasing fuel economy.",
    hjust = "left", color = "red"
  ) +
  annotate(
    geom = "segment",
    x = 3, y = 35, xend = 5, yend = 25, color = "blue",
    arrow = arrow(type = "closed")
  )
```

3.  Geom text puts a label at each point on every facet.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(vars(drv)) +
  geom_text(label = "a", nudge_y = 6)


ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~drv) +
  geom_text(data = mpg |> 
              filter(drv == "f"), aes(label = drv), nudge_x = 1)



ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~drv) +
  geom_text(data = mpg |> filter(drv == "f"), aes(label = drv), nudge_x = 1) +
   geom_text(data = mpg |> filter(drv == "4"), aes(label = drv), nudge_x = 1) +
   geom_text(data = mpg |> filter(drv == "r"), aes(label = drv), nudge_x = 1)
```

4\.

```{r}
ggplot(mpg,aes(x = hwy, y = displ, color = drv)) +
  geom_point() +
  geom_label(aes(label = drv))
```

5\.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  annotate(
    geom = "label", x = 3.5, y = 38,
    label = "Increasing engine size is \nrelated to decreasing fuel economy.",
    hjust = "left", color = "red"
  ) +
  annotate(
    geom = "segment",
    x = 3, y = 35, xend = 5, yend = 25, color = "blue",
    arrow = arrow(type = "closed")
  )

```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  annotate(
    geom = "label", x = 3.5, y = 38,
    label = "Increasing engine size is \nrelated to decreasing fuel economy.",
    hjust = "left", color = "red"
  ) +
  annotate(
    geom = "segment",
    x = 3, y = 35, xend = 5, yend = 25, color = "blue",
    arrow = arrow(type = "open")
  )
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  annotate(
    geom = "label", x = 3.5, y = 38,
    label = "Increasing engine size is \nrelated to decreasing fuel economy.",
    hjust = "left", color = "red"
  ) +
  annotate(
    geom = "segment",
    x = 3, y = 35, xend = 5, yend = 25, color = "blue",
    arrow = arrow(ends = "first",type = "closed")
  )
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  annotate(
    geom = "label", x = 3.5, y = 38,
    label = "Increasing engine size is \nrelated to decreasing fuel economy.",
    hjust = "left", color = "red"
  ) +
  annotate(
    geom = "segment",
    x = 3, y = 35, xend = 5, yend = 25, color = "blue",
    arrow = arrow(length = unit(0.70, "inches"),type = "closed")
  )
```

## 12.4.6 Exercises {.unnumbered}

1.  Because geom hex already fills this in you could use scale_fill instead.

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_color_gradient(low = "black", high = "red") +
  coord_fixed()
```

2.  Name or palette ,you can label your x and y axis with scale_x_continous and also labs.

3.  

```{r}
presidential |>
  mutate(id = 33 + row_number()) |>
  ggplot(aes(x = start, y = id, color = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_color_manual(values = c(Republican = "#E81B23", Democratic = "#00AEF3")) +
    scale_x_date(name = "Term", breaks = seq(from = ymd("1953-01-20"), to = ymd("2021-01-20"), by = "4 years"), date_labels = "'%y") +
scale_y_continuous(
  name = "president",
  breaks = 34:45,
  labels = presidential$name
)
```

4.  

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(color = cut), alpha = 1/20) +
    guides(color = guide_legend(override.aes = list(alpha = 1) ) )
```

## 12.5.1 Exercises {.unnumbered}

1.  

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(color = cut), alpha = 1/20) +
    guides(color = guide_legend(override.aes = list(alpha = 1) ) ) +
  theme_economist()
```

2\.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() + 
xlab(substitute(paste(bold("something")))) +
  ylab(substitute(paste(bold("something"))))

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  theme(axis.title = element_text(color = "blue",face = "bold"))
```

## 12.6.1 Exercises {.unnumbered}

1.It looks like it turns the plots like vertical if that makes sense.

```{r}
p1 <- ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  labs(title = "Plot 1")
p2 <- ggplot(mpg, aes(x = drv, y = hwy)) + 
  geom_boxplot() + 
  labs(title = "Plot 2")
p3 <- ggplot(mpg, aes(x = cty, y = hwy)) + 
  geom_point() + 
  labs(title = "Plot 3")

(p1 | p2) / p3
```

2\.

```{r}
p3 / (p2 | p1)
```
