---
title: "Communication"
---

## Prerequisites {.unnumbered}

```{r}
#| label: setup

library(tidyverse)   # the tidyverse
library(scales)      # to adjust display of numbers
library(ggrepel)     # to clearly position text labels
library(patchwork)   # to display multiple plots
library(gt)          # to display beautiful tables in Quarto
```

## 12.2.1 Exercises {.unnumbered}

1.  The plot is shown below in @fig-q1-ex2.

```{r}
#| label: fig-q1-ex2
#| fig-cap: "Scatter plot of highway mileage vs. engine displacement for different class of vehicles"


data(mpg)
mpg |>
  ggplot(aes(x = displ,
             y = hwy,
             col = as_factor(class)
             )
         ) +
  geom_point() +
  labs(title = "Higher displacement vehicles have lower mileage",
       subtitle = "2-seater sports cars have high mileage despite higher displacements",
       x = "Engine Displacement (in liters)",
       y = "Highway Mileage (miles per gallon)",
       caption = "Data from mpg dataset",
       color = "Vehicle Class") +
  theme(legend.position = "right") +
  scale_color_viridis_d()
```

2.  The code is shown below:
```{r}
#| label: fig-q2-ex2
#| fig-cap: "A scatter-plot of highway vs city MPG to recreate the plot shown in question"

mpg |>
  ggplot(aes(x = cty,
             y = hwy,
             shape = drv,
             col = drv)) +
  geom_point() +
  labs(x = "City MPG", y = "Highway MPG",
       col = "Type of \ndrive train",
       shape = "Type of \ndrive train")
```

3.  I use a figure produced for Question 5 of Exercise 11.5.3.1 to demonstrate outliers in the `diamonds` dataset, based on a combination of `x` and `y` attributes of various diamonds. I use the `ggrepel` package[@ggrepel] to label the outliers.

```{r}
#| label: fid-q3-ex2
#| fig-cap: "A scatter plot of x and y dimensions of a diamond to detect outliers"
#| fig-cap-location: top

diamonds |>    
    mutate(res_lm = (lm(diamonds$y ~ diamonds$x)$residuals)) |>
    filter(x >= 4) |>
    mutate(res_lm = res_lm < -1 | res_lm > 1) |>
    rowid_to_column() |>
    mutate(rowid = ifelse(res_lm, rowid, NA)) |>
    ggplot(aes(x = x, 
               y = y,
               label = scales::comma(rowid))) +   
    geom_point(alpha = 0.2) +
    coord_cartesian(xlim = c(4, 11), ylim = c(4, 11)) +
    ggrepel::geom_text_repel(size = 3) +
    labs(x = "Length of diamond, in mm (x)", 
         y = "Width of diamond, in mm (y)",
         title = "Certain diamonds have an abnormal shape",
         subtitle = "The labelled diamonds, with their IDs shown, have excessive length or width",
         caption = "Data from diamonds data-set, ggplot2") +
  theme_light()
```

## 12.3.1 Exercises {.unnumbered}

1.  The @fig-q1-ex3 show the plot produced. The important thing to understand is the use of arguments `hjust` and `vjust` to the function `geom_text()` : --

  A.  **`hjust` (Horizontal Justification):**

    -   The `hjust` argument controls the horizontal alignment or justification of the text relative to its specified x-coordinate.

    -   It takes values between 0 and 1, where 0 means left-aligned, 1 means right-aligned, and 0.5 means centered.

    -   Values less than 0 or greater than 1 can be used to align the text outside the range of the plot area.

    -   Negative values shift the text label to the left, and values greater than 1 shift it to the right.

  B.  **`vjust` (Vertical Justification):**

    -   The `vjust` argument controls the vertical alignment or justification of the text relative to its specified y-coordinate.

    -   Like `hjust`, it takes values between 0 and 1, where 0 means bottom-aligned, 1 means top-aligned, and 0.5 means centered.

    -   Values less than 0 or greater than 1 can be used to align the text outside the range of the plot area.

    -   Negative values shift the text label downward, and values greater than 1 shift it upward.

```{r}
#| label: fig-q1-ex3
g = mpg |>
  ggplot(aes(x = displ,
             y = hwy)) +
  geom_point(position = position_jitter(seed = 1)) +
  theme_light()
  
g +
  # Add text labels at the four corners using infinite positions
  geom_text(aes(x = Inf, y = Inf, 
                label = "High mileage, high displacement"),
            hjust = 1, vjust = 1) +
  geom_text(aes(x = -Inf, y = -Inf, 
                label = "Low mileage, low displacement"),
            hjust = 0, vjust = -0.2) +
  geom_text(aes(x = Inf, y = -Inf, 
                label = "Low mileage, high displacement"),
            hjust = 1, vjust = -0.2) +
  geom_text(aes(x = -Inf, y = Inf, 
                label = "High mileage, Low displacement"),
            hjust = 0, vjust = 1)
```

2.  
```{r}
# Compute middle of x and y axis
x_mid = (min(mpg$displ) + max(mpg$displ))/2
y_mid = (min(mpg$hwy) + max(mpg$hwy))/2
lbl_grph = str_wrap("This is a point in the middle of the previous graph",
                    width = 30)

g +
  annotate(
    geom = "point",
    x = x_mid,
    y = y_mid,
    shape = 23,
    size = 6,
    color = "red",
    fill = "pink"
  ) +
  annotate(
    geom = "text",
    label = lbl_grph,
    x = x_mid,
    y = y_mid,
    color = "red",
    hjust = 0,
    vjust = -0.5,
    size = 6
  )
```

3.  In `ggplot2`, the `geom_text()` function is used to add text labels to a plot. Labels added using `geom_text()` can interact with faceting. Faceting involves creating multiple plots based on different subsets of the data. The following points clarify the interaction methods: --

**Interaction between `geom_text()` labels and faceting:** When we use faceting in ggplot2 (with functions like `facet_wrap()` or `facet_grid()`), the labels created by `geom_text()` are typically placed in each facet based on the data corresponding to that specific facet. This means that the labels are positioned within each facet's coordinate system. @fig-q3a-ex3 is an example using `gtcars` dataset from `gt` package to display the name (`geom_text`) of *most fuel efficient car on higway within each body style*: --

```{r}
#| label: fig-q3a-ex3
#| fig-cap: "Scatter-plot of highway mileage vs. horsepower, for different body type of cars - with most fuel efficient car in each categry labelled"
#| code-fold: true

data("gtcars")

gtcars |>
  # Remove NAs, else they will be treated as maximum 
  drop_na() |>
  
  # Cosmetic improvements for labelling in plot
  mutate(bdy_style = str_to_title(bdy_style)) |>
  mutate(drivetrain = case_when(
    drivetrain == "awd" ~ "All-Wheel Drive",
    drivetrain == "rwd" ~ "Rear-Wheel Drive"
  )) |>
  
  # Determine car with maximum Highway mileage in each group
  group_by(bdy_style) |>
  mutate(
    high_cat = ifelse(mpg_h == max(mpg_h),
                      yes = paste0(mfr, " ", model),
                      no = NA)
    ) |>
  ggplot(aes(x = hp,
             y = mpg_h,
             color = drivetrain,
             label = high_cat)) +
  geom_point(size = 1.5) +
  geom_text_repel(color = "#454647") +
  
  # To makr mpg_h comparable visually across facets,
  # we select layout of 4 columns
  facet_wrap(~bdy_style, ncol = 4) +
  
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Horsepower", 
       y = "Highway Fuel efficiency (miles per gallon)",
       color = "Car's drivetrain",
       title = "Most fuel efficient car in each category has the least HP",
       caption = "Data grom gt package (gtcars)")
```

**Adding a label to a single facet:** If you want to add a label to only one specific facet, you can do so by adding a new data frame containing the label information for that facet; or, by creating a new column with the value stored only for the observation to be labelled, and rest as `NA`s. @fig-q3b-ex3 is an example using `gtcars` dataset from `gt` package to display the name (`geom_text`) of *most fuel efficient car on the highway in the entire data set*: --

```{r}
#| label: fig-q3b-ex3
#| fig-cap: "Scatter-plot of highway mileage vs. horsepower, for different body type of cars - with most fuel efficient car in dataset labelled"
#| code-fold: true

gtcars |>
  # Remove NAs, else they will be treated as maximum 
  drop_na() |>
  
  # Cosmetic improvements for labelling in plot
  mutate(bdy_style = str_to_title(bdy_style)) |>
  mutate(drivetrain = case_when(
    drivetrain == "awd" ~ "All-Wheel Drive",
    drivetrain == "rwd" ~ "Rear-Wheel Drive"
  )) |>
  
  # Determine car with maximum Highway mileage (so, no need to Group)
  # group_by(bdy_style) |>
  mutate(
    high_cat = ifelse(mpg_h == max(mpg_h),
                      yes = paste0(mfr, " ", model),
                      no = NA)
    ) |>
  ggplot(aes(x = hp,
             y = mpg_h,
             color = drivetrain,
             label = high_cat)) +
  geom_point(size = 1.5) +
  geom_text_repel(color = "#454647") +
  
  # To makr mpg_h comparable visually across facets,
  # we select layout of 4 columns
  facet_wrap(~bdy_style, ncol = 4) +
  
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Horsepower", 
       y = "Highway Fuel efficiency (miles per gallon)",
       color = "Car's drivetrain",
       title = "Most fuel efficient car in the data-set is BMW 6-Series",
       caption = "Data grom gt package (gtcars)")

```

**Putting a different label in each facet:** The easiest way to annotate for each facet separately is to create a dataframe with a row for each facet. For example, in @fig-q3c-ex3, I create the same plot as above, with an annotation in each facet of the plot, highlighting my main message for that body style. I create these annotations using `geom_label()`. I could do the same with `geom_text()` as well.

```{r}
#| label: fig-q3c-ex3
#| fig-cap: "Scatter-plot of highway mileage vs. horsepower, for different body type of cars - with different annotations in each facet"
#| code-fold: true
#| fig-width: 8


annt_con = "Convertibles generally have lower HP"
annt_cup = "Fuel efficiency decreases with rising HP"
annt_sed = "Sedans generally have lower HP"
annt_hbk = "Data-set has only two hatchbacks"

annt_df = tibble(
  bdy_style = as_vector(distinct(gtcars, bdy_style)),
  annt = str_wrap(c(annt_con, annt_cup, annt_sed, annt_hbk), 
                  width = 15),
  x_annt = 800,
  y_annt = 27.5
  ) |>
  mutate(bdy_style = str_to_title(bdy_style))

gtcars |>
  
  # Cosmetic improvements for labelling in plot
  mutate(bdy_style = str_to_title(bdy_style)) |>
  mutate(drivetrain = case_when(
    drivetrain == "awd" ~ "All-Wheel Drive",
    drivetrain == "rwd" ~ "Rear-Wheel Drive"
  )) |>
  
  ggplot(aes(x = hp,
             y = mpg_h,
             color = drivetrain)) +
  geom_point(size = 1.5) +
  
  # To make mpg_h comparable visually across facets,
  # we select layout of 4 columns
  facet_wrap(~bdy_style, ncol = 4) +
  geom_label(
    data = annt_df,
    mapping = aes(x = x_annt,
                  y = y_annt,
                  label = annt),
    color = "#454647",
    fontface = "bold",
    label.size = 0.1,
    size = 3
  ) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Horsepower", 
       y = "Highway Fuel efficiency (miles per gallon)",
       color = "Car's drivetrain",
       title = "Higher HP lowers highway mileage",
       caption = "Data grom gt package (gtcars)")
```

4.  In ggplot2, the `geom_label()` function is used to add labeled annotations to a plot, as shown above in @fig-q3c-ex3. To control the appearance of the background box of these labels, we can use the following arguments:

    1.  `label.padding`: This argument controls the padding around the label text within the background box, with a default value of 0.25.

    2.  `label.size`: This argument defines the size of the label border, in mm.

    3.  `label.r`: This argument sets the radius of the rounded corners of the background box, with a default value of 0.15.

    An example of how we can use these arguments in the `geom_label()` function is shown below in :

```{r}
#| label: fig-q4-ex3
#| fig-cap: "A simple scatterplot with background of the label changed"
#| fig-cap-location: top

tibble(x = 1:10, 
       y = 2*x + rnorm(10)) |>
  
  ggplot(aes(x = x, y = y)) +
  geom_label(aes(label = x), 
             label.padding = unit(1, "lines"),  # Adjust padding
             label.size = 2,                    # Set text size
             label.r = unit(1, "lines")) +      # Set corner radius
  theme_void()

```

5.  In `ggplot2` and `R`, the `arrow()` function is used to customize the appearance of arrowheads in line segments. As the results show in @fig-q5-ex3, the `arrow()` function has four main arguments that we can use to control the arrowheads:

    1.  **`type`:** This argument specifies the type of arrowhead. There are three main types:

        -   `"closed"`: A closed arrowhead with a triangular shape.

        -   `"open"`: An open arrowhead with a triangular shape.

    2.  **`length`:** This argument controls the length of the arrowhead. We can specify it using a numeric value or a `unit()` object, which allows you to define the length in various units such as `"inches"`, `"cm"`, etc.

    3.  **`ends`**: This argument determines which end of the line segment the arrowhead should appear on. It can take two values:

        -   `"last"`: Arrowhead is placed at the end of the segment (default behavior).

        -   `"first"`: Arrowhead is placed at the start of the segment.

        -   `"both"`: Arrowhead appears at both the start and end of the segment.

    4.  **`angle`**: This argument controls the angle of the arrow head in degrees *(smaller numbers produce narrower, pointier arrows)*. It describes the width of the arrow head.

```{r}
#| label: fig-q5-ex3
#| fig-cap: "Customizing the arrows in geom_segment() using the 4 main arguments to arrow()"
#| fig-cap-location: top
#| fig-asp: 1
  
nos = 8                                     # Number of spokes to create
angle = seq(0, 7/4 * pi, length.out = nos)  # Angles for 8 directions
arrow_length = 0.8                          # Length of line segment

g = tibble(
    x_start = rep(0, nos),              # Common starting point
    y_start = rep(0, nos),
    x_end = cos(angle) * arrow_length,  # Calculate end points based on angles
    y_end = sin(angle) * arrow_length
    ) |>
    ggplot(aes(x = x_start,
               y = y_start,
               xend = x_end,
               yend = y_end)) +
  theme_void() 

gridExtra::grid.arrange(
  
  g + geom_segment(arrow = arrow(type = "open")) + 
    labs(subtitle = "arrow(type = open)"),

  g + geom_segment(arrow = arrow(type = "closed")) + 
    labs(subtitle = "arrow(type = closed)"),  
  
  g + geom_segment(arrow = arrow(angle = 15)) + 
    labs(subtitle = "arrow(angle = 15)"),

  g + geom_segment(arrow = arrow(angle = 90)) + 
    labs(subtitle = "arrow(angle = 90)"),  
    
  g + geom_segment(arrow = arrow(ends = "first")) + 
    labs(subtitle = "arrow(ends = first)"),

  g + geom_segment(arrow = arrow(ends = "last")) + 
    labs(subtitle = "arrow(ends = last)"),

  g + geom_segment(arrow = arrow(ends = "both")) + 
    labs(subtitle = "arrow(ends = both)"),
  
  g + geom_segment(arrow = arrow(length = unit(1, "cm"))) + 
    labs(subtitle = "arrow(length = unit(1, cm)"),
  
  g + geom_segment(arrow = arrow(length = unit(0.5, "cm"))) + 
    labs(subtitle = "arrow(length = unit(0.5, cm)"),
  
  ncol = 3
)
```

## 12.4.6 Exercises {.unnumbered}

1.  In the code provided, we're using the `geom_hex()` function to create a hexbin plot. However, the `scale_color_gradient()` function is used to adjust the color scale of a continuous color aesthetic.

In present case, the `geom_hex()` function uses the **`fill` aesthetic** to represent the density of points in each hexagonal bin, but we're trying to adjust the color scale using `scale_color_gradient()`. So, to adjust the color scale of the hex-bin plot, we should use the **`scale_fill_gradient()`** function. Here's the corrected version of the code:

```{r}
#| code-fold: true

df <- tibble(   
  x = rnorm(10000),   
  y = rnorm(10000) 
  )  

ggplot(df, aes(x, y)) +   
  geom_hex() +   
  scale_fill_gradient(low = "white", 
                      high = "red") +   
  coord_fixed()
```

2.  The first argument to every scale function is `name` . It is used as the axis or legend title.

    -   If `name = waiver()`, the default, then the name of the scale is taken from the first mapping used for that aesthetic.

    -   If `name = NULL`, then then legend title will be omitted.

The `labs()` can also be used to name a legend, with use the aesthetic name to define the new name. For example, `labs(x = NULL)`. The output will be the same.

However, when we use `scale_` functions, we are customizing how a specific aesthetic is represented in the plot. It allows us to modify the scale, transformation, and appearance of the aesthetic. On the other hand, `labs()`is specifically designed for modifying the text labels and titles of different plot components. It's not concerned with how an aesthetic is represented or scaled in the plot.

3.  We can change the display of the presidential terms by:

a.  **Combining the two variants that customize colors and x axis breaks.**

b.  **Improving the display of the y axis.**

c.  **Labeling each term with the name of the president.**

d.  **Adding informative plot labels.**

I have attempted to answer the first four of these sub-questions simultaneously, in the code below. The resulting plot is at @fig-q3a-ex4 .

```{r}
#| label: fig-q3a-ex4
#| fig-cap: "Changing the display of the presidential terms by customizing colours, axes and labels"
#| fig-width: 10

data("presidential")

# Creating a vector of years where a new president takes office
# to allow plotting it on the x-axis
x_labs = 
  presidential |>
  select(start) |>
  as_vector() |>
  as_date()

y_labs = seq(
  from = 1 + 33,
  to = nrow(presidential) + 33,
  by = 1
)

g = 
presidential |>
  as_tibble() |>
  mutate(id = 33 + row_number()) |>
  ggplot(aes(x = start, xend = end, 
             y = id, yend = id, 
             col = party,
             label = name)) +
  geom_segment(lwd = 10) +
  
  # (a) Combining the two variants that customize colors and x axis breaks.
  scale_color_manual(values = c("Democratic" = "blue",
                                "Republican" = "red")) +
  scale_x_continuous(breaks = x_labs,
                     labels = year(x_labs)) +
  
  
  # (b) Improving the display of the y axis.
    scale_y_reverse(breaks = y_labs,
                  labels = y_labs) +
  
  
  # (c) Labeling each term with the name of the president.
  geom_text(hjust = -0.05,
            vjust = -1.5) +
  
  geom_rect(aes(xmin = start,
                xmax = end,
                ymin = 33.5,
                ymax = 45.5,
                fill = party),
            alpha = 0.3,
            col = NA) +
  scale_fill_manual(values = c("Democratic" = "blue",
                                "Republican" = "red")) +
  
  
  # (d) Adding informative plot labels.
  labs(color = "President's Political Party",
       y = "Table of Precedence Number", 
       x = "Year") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(panel.grid.major.x = element_line(color = "darkgrey"),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.y = element_text()) +
  guides(fill = "none")

print(g) 
```

    e.  Placing breaks every 4 years (this is trickier than it seems!).

To place breaks after every four years, I create a new vector of dates, called `x_4yr_labs` and then use it, as shown below with `scale_x_continuous()`. The resulting figure is at @fig-q3b-ex4.

```{r}
#| label: fig-q3b-ex4
#| fig-cap: "Customizing plots by placing breaks every 4 years"
#| code-fold: true
#| fig-width: 10

x_4yr_labs = 
  presidential |>
  select(start) |>
  as_vector() |>
  as_date()
  
x_4yr_labs = seq(from = ymd(x_4yr_labs[1]),
                 to = ymd(x_4yr_labs[length(x_4yr_labs)]),
                 by = "4 years")

g + 
  scale_x_continuous(breaks = x_4yr_labs,
                     labels = year(x_4yr_labs))
```

### Question 4

**First, create the following plot. Then, modify the code using `override.aes` to make the legend easier to see.**

```         
ggplot(diamonds, 
       aes(x = carat, y = price)) +   
  geom_point(aes(color = cut), 
             alpha = 1/20)
```

In the code below, I've improved the original plot. I used some extra code: `guides(color = guide_legend(override.aes = list(size = 5, alpha = 0.5)))`. This changes how the point marks look in the legend, making them a bit darker so the legend colors are easier to notice. Also, I made the points a bit bigger to help see them better. You can see the final plot in @fig-q4-ex4. I also changed the background to white so that the transaprent points are clearer.

```{r}
#| label: fig-q4-ex4
#| fig-cap: "Modifying the code using override.aes to make the legend easier to see"

ggplot(diamonds, aes(x = carat, y = price)) +   
  geom_point(aes(color = cut), alpha = 1/20) +
  theme_light() +
  guides(color = guide_legend(override.aes = list(size = 5,
                                                  alpha = 0.5)))
```

## 12.5.1 Exercises {.unnumbered}

1.  Here in @fig-q1-ex5, I use the `theme_wsj()` (Theme based on the plots in *The Wall Street Journal*) offered by `ggthemes` package[@ggthemes] and apply it to the plot made above in @fig-q4-ex4.

```{r}
#| label: fig-q1-ex5
#| fig-cap: "Using the theme_wsj() offered by ggthemes package"
#| fig-cap-location: top

ggplot(diamonds, aes(x = carat, y = price)) +   
  geom_point(aes(color = cut), alpha = 1/20) +
  theme_light() +
  guides(color = guide_legend(override.aes = list(size = 5,
                                                  alpha = 0.5))) +
  ggthemes::theme_wsj()
```

2.  I do so by adding the following code to my previous plot: `theme(axis.text = element_text(face = "bold", color = "blue", size = 6))` . The resulting plot is displayed in @fig-q2-ex5 .

```{r}
#| label: fig-q2-ex5
#| fig-cap: "Figure make the axis labels blue and bold."
#| fig-cap-location: top

ggplot(diamonds, aes(x = carat, y = price)) +   
  geom_point(aes(color = cut), alpha = 1/20) +
  theme_light() +
  guides(color = guide_legend(override.aes = list(size = 5,
                                                  alpha = 0.5))) +
  theme(axis.text = element_text(face = "bold",
                                 color = "blue",
                                 size = 12)) +
  theme(legend.position = "bottom")
```

## 12.6.1 Exercises {.unnumbered}

1.  
```{r}
#| label: fig-q1-ex6
#| fig-cap: "Plot produced by the code given in the question"

p1 <- ggplot(mpg, aes(x = displ, y = hwy)) +    
    geom_point() +    
    labs(title = "Plot 1") 

p2 <- ggplot(mpg, aes(x = drv, y = hwy)) +    
    geom_boxplot() +    
    labs(title = "Plot 2") 
    
p3 <- ggplot(mpg, aes(x = cty, y = hwy)) +    
    geom_point() +    
    labs(title = "Plot 3")  

(p1 | p2) / p3
```

In `patchwork`, the pipe operator `|` is used to arrange plots side by side, and the `/` operator is used to arrange plots vertically. Thus, with parenthesis, as shown above in @fig-q1-ex6 .

  1.  `(p1 | p2)` arranges plots `p1` and `p2` side by side because of the `|` operator.

  2.  The result of `(p1 | p2)` (a combined plot of `p1` and `p2`) is then combined with plot `p3` vertically using the `/` operator.

  **Now, when we omit the parentheses,** as we can see in @fig-q1a-ex6 ,

  1.  `p2 / p3` is evaluated first because `/` has a higher precedence than `|`.

  2.  `p2 / p3` arranges plots `p2` and `p3` vertically because of the `/` operator.

  3.  The result of `p2 / p3` is then combined with plot `p1` using the `|` operator.

This is because the `/` has a higher precedence than `|` , in other words, `R` applies ***BODMAS rules (or, PEMDAS)***[^1] to take the division operator `/` before the `|` (pipe) operator, and we get Plot 2 above Plot 3, while Plot 1 stays in a separate column. This happens perhaps because division comes before addition in normal mathematics, and the `patchwork` creators have followed that convention.

[^1]:
    > Note: The BODMAS rules (also known as PEMDAS or PEDMAS in different regions) are a set of rules that dictate the order of operations to be followed when evaluating mathematical expressions. These rules ensure that mathematical expressions are evaluated consistently and accurately. BODMAS stands for:
    >
    > 1.  **B**rackets: Evaluate expressions inside brackets first.
    >
    > 2.  **O**rders: Evaluate exponential or power operations.
    >
    > 3.  **D**ivision and **M**ultiplication: Perform division and multiplication.
    >
    > 4.  **A**ddition and **S**ubtraction: Perform addition and subtraction operations from left to right.

```{r}
#| label: fig-q1a-ex6
#| fig-cap: "Plot produced by the code when the parenthesis are omitted"

p1 <- ggplot(mpg, aes(x = displ, y = hwy)) +    
    geom_point() +    
    labs(title = "Plot 1") 

p2 <- ggplot(mpg, aes(x = drv, y = hwy)) +    
    geom_boxplot() +    
    labs(title = "Plot 2") 
    
p3 <- ggplot(mpg, aes(x = cty, y = hwy)) +    
    geom_point() +    
    labs(title = "Plot 3")  

p1 | p2 / p3
```

2.  Using `patchwork`[@patchwork] is simple, and we can create @fig-q2-ex6 by the code shown below: --

```{r}
#| label: fig-q2-ex6
#| fig-cap-location: top
#| fig-cap: "Recreating the plot given in Question 2, Exerise 12.6.1 using the patchwork package"

p1 / (p2 | p3) +
  plot_annotation(tag_levels = 'A',
                  tag_prefix = "Fig. ")
```

