---
title: "Data tidying"
---

```{r}
#| results: "asis"
#| echo: false

source("_common.R")
```

## Prerequisites {.unnumbered}

```{r}
library(tidyverse)
```

## 6.2.1 Exercises {.unnumbered}

1.  In each of `table1`, `table2`, and `table3`, each observation represents a country.
    In `table1`, `country` is the country name, `year` is the year of data collection, `cases` is the number of people with the disease in that year, and `population` is the number of people in each country in that year.
    In `table2`, `country` and `year` are the same as in `table1`, `type` is the type of number, and `count` is the number of observations (either `cases` or `population` depending on `type`).
    Finally, in `table3`, `country` and `year` are again the same as in `table1`, and `rate` is the rate of disease (cases divided by population).

2.  For `table2`, we need to reshape the data have a column for cases and a column for population and then divide the two to calculate the rate.
    A possible approach is shown below.

    ```{r}
    table2 |>
      pivot_wider(
        names_from = type,
        values_from = count
      ) |> 
      mutate(rate = cases / population * 10000)
    ```

    For `table3`, we need to separate cases and population into their own columns and then divide them.
    A possible approach is shown below.

    ```{r}
    table3 |>
      separate_wider_delim(
        cols = rate, 
        delim = "/", 
        names = c("cases", "population"),
      ) |>
      mutate(
        cases = as.numeric(cases),
        population = as.numeric(population),
        rate = cases / population * 10000
      )
    ```
