---
title: "Data transformation"
---

```{r}
#| results: "asis"
#| echo: false

source("_common.R")
```

## Prerequisites {.unnumbered}

```{r}
library(nycflights13)
library(tidyverse)
```

## 4.2.5 Exercises {.unnumbered}

1.  Pipelines for each part are given below.

<!-- -->

a.  

```{r}
flights |>
  filter(arr_delay >= 120) |>
  arrange(desc(arr_delay))
```

b.  

```{r}
flights |>
  filter(dest %in% c("IAH", "HOU"))
```

c.  

```{r}
flights |>
  filter(carrier %in% c("UA", "DL", "AA"))
```

d.  

```{r}
flights |>
  filter(month %in% c(7, 8, 9))
```

e.  

```{r}
flights |> 
  filter(arr_delay >= 120 & dep_delay <= 0) |> view()
```

f.  

```{r}
flights |> 
  filter(dep_delay >= 60 & dep_delay - arr_delay > 30)
```

2.  Flights with longest departure delays and, among those, flights that left earliest in the morning:

```{r}
flights |> 
  arrange(desc(dep_delay)) |> 
  arrange(sched_dep_time) |>
  relocate(dep_delay, sched_dep_time)
```

3.  Fastest flights, measured as miles per hour:

```{r}
flights |> 
  mutate(speed = distance / (air_time / 60)) |>
  arrange(desc(speed)) |>
  relocate(speed)
```

4.  Yes, there was a flight on every day of 2013 since there are 365 distinct combinations of `year`, `month`, and `day`, which is equal to the number of days in the year 2013.

```{r}
flights |> 
  distinct(year, month, day) |>
  nrow()
```

5.  Flights that traveled the farthest distance:

```{r}
flights |> 
  arrange(desc(distance)) |>
  relocate(distance)
```

Flights that travelled the shortest distance

```{r}
flights |> 
  arrange(distance) |>
  relocate(distance)
```

6.  The order doesn't matter because we filter based on a condition, not based on row number.

## 4.3.5 Exercises {.unnumbered}

1.  I would expect `dep_time` to be `sched_dep_time + dep_delay`.

```{r}
flights |> 
  relocate(dep_time, sched_dep_time, dep_delay)
```

2.  The following are some of the ways these variables can be selected.

```{r}
flights |> 
  select(dep_time, dep_delay, arr_time, arr_delay)

flights |> 
  select(starts_with("dep"), starts_with("arr"))

flights |>
  select(dep_time:arr_delay, -contains("sched"))
```

3.  You get the variable just once.

```{r}
flights |> 
  select(dep_time, dep_time)
```

4.  You ask if `any_of()` these variables have a certain thing you are looking for.

```{r}
variables <- c("year", "month", "day", "dep_delay", "arr_delay")

flights |> 
  select(any_of(variables))
```

5.  Yes, it does surprise me since the variable names are lowercase but the string in `contains()` is uppercase. By default, `contains()` ignores case.

```{r}
flights |> 
  select(contains("TIME"))
```

To change this default behavior, set `ignore.case = FALSE`.

```{r}
flights |> 
  select(contains("TIME", ignore.case = FALSE))
```

6.  Below we rename `air_time` to `air_time_min` and move it to the beginning of the data frame.

```{r}
flights |>
  rename(air_time_min = air_time) |>
  relocate(air_time_min)
```

7.  This doesn't work because the result of the `select()` step is a data frame with only the `tailnum` variable, so it's not possible to arrange it by another variable, `arr_delay`.

```{r}
#| error: true

flights |> 
  select(tailnum) |> 
  arrange(arr_delay)
```

## 4.5.7 Exercises {.unnumbered}

1.  F9 (Frontier Airlines) has the worst average delays.

```{r}
flights |>
  group_by(carrier) |>
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) |>
  arrange(desc(avg_dep_delay))
```

2.  The following are the top 5 most departure delayed flights from each destination.

```{r}
flights |> 
  group_by(dest) |> 
  arrange(dest, desc(dep_delay)) |>
  slice_head(n = 5) |>
  relocate(dest, dep_delay)
```

3.  Over the course of the day, hourly average departure delay increases until about 7pm, and then declines again, however doesn't go as low as the beginning of the day.

```{r}
flights |>
  group_by(hour) |>
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) |>
  ggplot(aes(x = hour, y = avg_dep_delay)) + 
  geom_smooth()
```

4.  Supplying a negative value arranges the data frame in either ascending (with `slice_min()`) or descending (with `slice_max()`) order, but it doesn't actually slice the data frame for the lowest/highest values of the given variable.

```{r}
flights |> 
  slice_min(dep_delay, n = -5) |>
  relocate(dep_delay)

flights |> 
  slice_min(dep_delay, n = 5) |>
  relocate(dep_delay)

flights |> 
  slice_max(dep_delay, n = -5) |>
  relocate(dep_delay)

flights |> 
  slice_max(dep_delay, n = 5) |>
  relocate(dep_delay)
```

5.  `count()` counts the number of observations in each group, setting the `sort` argument to `TRUE` arranges the categories in descending order of number of observations.

6.  First, let's define the data frame `df`.

```{r}
df <- tibble(
  x = 1:5,
  y = c("a", "b", "a", "a", "b"),
  z = c("K", "K", "L", "L", "K")
)
```

a.  The following groups `df` by `y`.

```{r}
df |>
  group_by(y)
```

b.  The following arranges `df` in ascending order of the value of `y`.

```{r}
df |>
  arrange(y)
```

c.  The following groups `df` by `y` and then calculates the average value of `x` for each group.

```{r}
df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
```

d.  The following groups `df` by `y` and `z`, and then calculates the average value of `x` for each group combination. The resulting data frame is grouped by `y`.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))
```

e.  The following groups `df` by `y` and `z`, and then calculates the average value of `x` for each group combination. The resulting data frame is not grouped.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x), .groups = "drop")
```

f.  Each of the following groups `df` by `y` and `z`, and then calculates the average value of `x` for each group combination. With `summarize()` the resulting data frame has one row per group combination while with `mutate()` the resulting data frame has the same number of rows as the original data frame.

```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
  group_by(y, z) |>
  mutate(mean_x = mean(x))
```
