---
title: "Factors"
execute: 
  echo: true
  warning: false
  error: false
  cache: true
---

## Prerequisites {.unnumbered}

```{r}
#| label: setup

library(tidyverse)
library(ggthemes)
library(gt)
library(gtExtras)
data("gss_cat")
```

## 17.3.1 Exercise

1.  The default bar chart is hard to understand because: --
    
    1.  It is vertical, and the names of categories overlap on x-axis.
    
    2.  The "Not applicable" category is before the lowest income group. Thus, the pattern is disturbed.
    
    We could improve the plot, as shown in @fig-q1-ex3, by: ---
    
    1.  Making it into a horizontal bar chart to allow space and easy reading of categories of income levels.
    
    2.  Move the "Not Applicable" level after the highest income level, along-side "Refused", "Dont' know" and "No answer".
    
    3.  Further, we could remove non-data ink, as per principles of Mr. Tufte to make our pattern stand out. Also, we could create a separate colouring scheme for data outside the income levels.
    
    ```{r}
    #| label: fig-q1-ex3
    #| fig-cap: "Improved bar chart"
    
    no_levels = levels(gss_cat$rincome)[c(1:3, 16)]
    
    gss_cat |>
      mutate(col_level = rincome %in% no_levels) |>
      ggplot(aes(y = fct_relevel(rincome, 
                                 "Not applicable",
                                 after = 3),
                 fill = col_level)) +
      geom_bar() +
      theme_minimal() +
      labs(x = "Number of respondents", y = NULL,
           title = "Income Levels of respondents in General Social Survey") +
      theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position = "none") +
      scale_fill_manual(values = c("#3d3b3b", "#999494"))
    ```
    
2.  The most common `relig` is "Protestant". And, the most common `partyid` is "Independent".
    
    ```{r}
    gss_cat |>
      count(relig, sort = TRUE)
    
    gss_cat |>
      count(partyid, sort = TRUE)
    ```


3.  We can see from the code below that more than one factor values in `denom` (denomination) occur only in "Protestant", "Christian" and "Other" religions. To explore further, we can cross-tabulate religion and denomination, as shown in @tbl-q3-ex3, and realize that the only religion to which denomination really applies to is "Protestant".
    
    We could also do a visualization as in @fig-q3a-ex3 .
    
    ```{r}
    gss_cat |>
      group_by(relig) |>
      summarise(n = n_distinct(denom)) |>
      arrange(desc(n)) |>
      filter(n > 1)
    ```
    
    ```{r}
    #| label: tbl-q3-ex3
    #| tbl-cap: "Cross-Table of the deominations within the three religions"
    
    gss_cat |>
      filter(relig %in% c("Protestant", "Christian", "Other")) |>
      group_by(relig, denom) |>
      tally() |>
      spread(relig, n) |>
      arrange(desc(Christian)) |>
      gt() |>
      sub_missing(missing_text = "") |>
      gt_theme_538()
    ```
    
    ```{r}
    #| label: fig-q3a-ex3
    #| tbl-cap: "Visualization of the number of deominations within religions"
    
    gss_cat |>
      group_by(relig) |>
      summarise(n = n_distinct(denom)) |>
      arrange(desc(n)) |>
      ggplot(aes(y = reorder(relig, n), x = n)) +
      geom_bar(stat = "identity") +
      theme_minimal() +
      labs(x = "Number of denominations", y = NULL,
           title = "Only Protestant religion has demoninations within it")
    ```

## 17.4.1 Exercises

1.  No, mean is not a good summary as the distribution of `tvhours` is right skewed. Instead, we should use median as a summary measure.
    
    ```{r}
    gss_cat |>
      drop_na() |>
      mutate(tvhours = as_factor(tvhours)) |>
      ggplot(aes(x = tvhours)) +
      geom_bar(col = "black", fill = "white") +
      theme_clean() +
      labs(x = "Hours per day spent watching TV",
           y = "Numbers", title = "Distribution of TV Hours is right skewed")
    ```

2.    **For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled.**
      
      The variables in the `gss_cat` data-set which are factors are: ---
      
      | Factor Variable | Levels                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | Order is arbitrary or principled                                                                                 |
      |-----------------|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------|
      | `marital`       | No answer, Never married, Separated, Divorced, Widowed and, Married                                                                                                                                                                                                                                                                                                                                                                                                                                                                              | Arbitrary, since they are not in a specific order                                                                |
      | `race`          | Other, Black, White and, Not applicable                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          | Arbitrary, since they are not in a specific order                                                                |
      | `rincome`       | No answer, Don't know, Refused, \$25000 or more, \$20000 - 24999, \$15000 - 19999, \$10000 - 14999, \$8000 to 9999, \$7000 to 7999, \$6000 to 6999, \$5000 to 5999, \$4000 to 4999, \$3000 to 3999, \$1000 to 2999, Lt \$1000 and, Not applicable                                                                                                                                                                                                                                                                                                | Principled, since the income levels are in a specified increasing or decreasing order, with few levels arbitrary |
      | `partyid`       | No answer, Don't know, Other party, Strong republican, Not str republican, Ind,near rep, Independent, Ind,near dem, Not str democrat and, Strong democrat                                                                                                                                                                                                                                                                                                                                                                                        | Partly Principled, as there are two extremes, and then levels in the middle.                                     |
      | `relig`         | No answer, Don't know, Inter-nondenominational, Native american, Christian, Orthodox-christian, Moslem/islam, Other eastern, Hinduism, Buddhism, Other, None, Jewish, Catholic, Protestant and, Not applicable                                                                                                                                                                                                                                                                                                                                   | Arbitrary, as the religions are not in a specific order.                                                         |
      | `denom`         | No answer, Don't know, No denomination, Other, Episcopal, Presbyterian-dk wh, Presbyterian, merged, Other presbyterian, United pres ch in us, Presbyterian c in us, Lutheran-dk which, Evangelical luth, Other lutheran, Wi evan luth synod, Lutheran-mo synod, Luth ch in america, Am lutheran, Methodist-dk which, Other methodist, United methodist, Afr meth ep zion, Afr meth episcopal, Baptist-dk which, Other baptists, Southern baptist, Nat bapt conv usa, Nat bapt conv of am, Am bapt ch in usa, Am baptist asso and, Not applicable | Arbitrary, as the denominations are not in a specific order.                                                     |

3.  Moving "Not applicable" to the front of the levels move it to the bottom of the plot, because `ggplot2` plots the levels in increasing order, starting bottom's upwards.

## 17.5.1 Exercises

1.  As reflected in the @fig-q1-ex5, the proportions of people identifying as Democrat has slightly increased, Republican has slightly decreased, and Independent has increased, over the period of 15 years reflected in the data-set.
    
    ```{r}
    #| label: fig-q1-ex5
    #| fig-cap: "Stacked bar chart of the partyid in the data-set"
    
    
    gss_cat |>
      mutate(
        partyid = fct_collapse(partyid,
          "Republican"  = c("Strong republican",  "Not str republican"),
          "Democrat"    = c("Strong democrat", "Not str democrat"),
          "Independent" = c("Independent", "Ind,near dem", "Ind,near rep"),
          "Others"      = c("No answer", "Don't know", "Other party")
        )
      ) |>
      group_by(year, partyid) |>
      count() |>
      ggplot(aes(x = year, y = n, fill = partyid)) +
      geom_col(position = "fill") +
      scale_fill_manual(values = c("lightgrey", "red", "grey", "blue")) +
      theme_classic() +
      theme(legend.position = "bottom") +
      labs(x = "Year", y = "Proportion of respondents", fill = "Party",
           subtitle = "Proportion of republicans has decreased, while that of independents has increased over the years",
           title = "In 15 years, share of parties' supporters has changed")
    ```

2.  We could collapse the `rincome` into a small set of categories using the following functions: --
    
    -   `fct_lump_n()`
    
    -   `fct_lump_lowfreq()`
    
    -   `fct_lump_min()`
    
    -   `fct_lump_prop()`
    
    -   `fct_lump()`
    
    -   `fct_collapse()`
    
    ```{r}
    gss_cat|>
      mutate(
        rincome = fct_lump_n(rincome, n = 6)
      ) |>
      group_by(rincome) |>
      count() |>
      arrange(desc(n)) |>
      ungroup() |>
      gt() |>
      cols_label(rincome = "Annual Income",
                 n = "Numbers")
    ```
3.    Yes, there are 9 groups (excluding other) in this example, as shown below also in @fig-q3-ex5. This is because `n = 10` argument limits the total groups to 10, and the function needs one group for ***"Other"***, i.e. all other groups whose count is lesser than top 9 groups. Thus, the groups shown are 9, with 1 as ***"Other"*** (at the end).
    
    ```{r}
    #| label: fig-q3-ex5
    #| fig-cap: "Table of number of respondents from each of top 10 religions, including Other"
    
    gss_cat |>
      mutate(relig = fct_lump_n(relig, n = 10)) |>
      count(relig) |>
      gt()
    ```
