---
title: "Hierarchial data"
---

### **Prerequisites**

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(repurrrsive)
library(jsonlite)
```

### **24.3.5 Exercises**

1.  What happens when you use [`unnest_wider()`](https://tidyr.tidyverse.org/reference/unnest_wider.html) with unnamed list-columns like `df2`?
    What argument is now necessary?
    What happens to missing values?

    When using `unnest_wider` with unnamed list-columns, we need to provide it with `names_sep` to generate automatic column names.

    ```{r}
    df2 <- tribble(
      ~x, ~y,
      1, list(11, 12, 13),
      2, list(21),
      3, list(31, 32),
    )

    df2 |> 
      unnest_wider(y, names_sep = "_")
    ```

2.  What happens when you use [`unnest_longer()`](https://tidyr.tidyverse.org/reference/unnest_longer.html) with named list-columns like `df1`?
    What additional information do you get in the output?
    How can you suppress that extra detail?

Each element get its own row and we get an additional "index" column specifying the inner names of these elements.

```{r}
  df1 <- tribble(
    ~x, ~y,
    1, list(a = 11, b = 12),
    2, list(a = 21, b = 22),
    3, list(a = 31, b = 32),
  )
  
  df1 |> 
    unnest_longer(y)
```

To suppress this extra detail, we can set the argument `indices_include` to `FALSE`

```{r}
  df1 |> 
    unnest_longer(y, indices_include = F)
```

2.  From time-to-time you encounter data frames with multiple list-columns with aligned values.
    For example, in the following data frame, the values of `y` and `z` are aligned (i.e. `y` and `z` will always have the same length within a row, and the first value of `y` corresponds to the first value of `z`).
    What happens if you apply two [`unnest_longer()`](https://tidyr.tidyverse.org/reference/unnest_longer.html) calls to this data frame?
    How can you preserve the relationship between `x` and `y`?
    (Hint: carefully read the docs).

    If we apply two `unnest_longer` to a dataframe, we get all the possible pairs of y and z per each x.

    ```{r}
    df4 <- tribble(
      ~x, ~y, ~z,
      "a", list("y-a-1", "y-a-2"), list("z-a-1", "z-a-2"),
      "b", list("y-b-1", "y-b-2", "y-b-3"), list("z-b-1", "z-b-2", "z-b-3")
    )

    df4 |>
      unnest_longer(y) |>
      unnest_longer(z)
    ```

    To preserve the relationship between x and y, we need to unnest simultaneously.

```{r}
df4 |>
  unnest_longer(c(y, z))
```

### **24.4.4 Exercises**

1.  Roughly estimate when `gh_repos` was created. Why can you only roughly estimate the date?

```{r}
repos <- tibble(gh_repos)

repos |>
  unnest_longer(gh_repos)|>
  unnest_wider(gh_repos) |>
  summarise(earliest_date = min(created_at), latest_date = max(created_at))
```

It probably wasn't created before June, 2012.
It's hard to tell when it was exactly created since we only have the creation date of individual repos and not the collection itself.

2.  The `owner` column of `gh_repo` contains a lot of duplicated information because each owner can have many repos.
    Can you construct an `owners` data frame that contains one row for each owner?
    (Hint: does [`distinct()`](https://dplyr.tidyverse.org/reference/distinct.html) work with `list-cols`?)

    ```{r}
    owners <- repos |>
      unnest_longer(gh_repos)|>
      unnest_wider(gh_repos)|>
      select(owner)|>
      unnest_wider(owner, names_sep = "_") |>
      distinct()

    owners
    ```

3.  Follow the steps used for `titles` to create similar tables for the aliases, allegiances, books, and TV series for the Game of Thrones characters.

    ```{r}
    chars <- tibble(json = got_chars)
    extract_got_table <- function(table_name){
      chars |> 
      unnest_wider(json) |> 
      select(id, name, table_name) |> 
      unnest_longer(table_name) |> 
      filter(c_across(table_name) != "")
      }

    aliases <- extract_got_table("aliases")
    allegiances <- extract_got_table("allegiances")
    books <- extract_got_table("books")
    tvSeries <- extract_got_table("tvSeries")

    books
    ```

    Explain the following code line-by-line.
    Why is it interesting?
    Why does it work for `got_chars` but might not work in general?

```{r}
tibble(json = got_chars) |> 
  unnest_wider(json) |> 
  select(id, where(is.list)) |> 
  pivot_longer(
    where(is.list), 
    names_to = "name", 
    values_to = "value"
  ) |>  
  unnest_longer(value)
```

At first we turned the list into list-columns to suit our tools.
After that, we unnested the named list-column wider to separate columns.
Then, we picked only the list-columns and the id column.
We used `pivot_longer` to lengthen the data, decreasing the number of columns.
Lastly, we unnested the last list-column.

This sequence of steps is unique to this particular data because it follows its structure.
A different list would have a different structure and would require a different sequence to "rectangle" it.

5.  In `gmaps_cities`, what does `address_components` contain? Why does the length vary between rows? Unnest it appropriately to figure it out. (Hint: `types` always appears to contain two elements. Does [`unnest_wider()`](https://tidyr.tidyverse.org/reference/unnest_wider.html) make it easier to work with than [`unnest_longer()`](https://tidyr.tidyverse.org/reference/unnest_longer.html)?) .

    ```{r}
    locations <- gmaps_cities |> 
      unnest_wider(json) |> 
      select(-status) |> 
      unnest_longer(results) |> 
      unnest_wider(results) 
    locations |>
      select(city, address_components) |>
      unnest_longer(address_components) |>
      unnest_wider(address_components) |>
      unnest_longer(types)

    ```

### **24.5.4 Exercises**

1.  Rectangle the `df_col` and `df_row` below. They represent the two ways of encoding a data frame in JSON.

```{r}
json_col <- parse_json('
  {
    "x": ["a", "x", "z"],
    "y": [10, null, 3]
  }
')
json_row <- parse_json('
  [
    {"x": "a", "y": 10},
    {"x": "x", "y": null},
    {"x": "z", "y": 3}
  ]
')

df_col <- tibble(json = list(json_col))
df_row <- tibble(json = json_row)

df_col |>
  unnest_wider(json)|>
  unnest_longer(c(x,y))

df_row |>
  unnest_wider(json)
```
