---
title: "Workflow: basics"
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
```

## 3.5 Exercises {.unnumbered}

1.  The variable is called `my_variable` while the next line calls it `my_varıable`, with an an `ı` (`i` without a dot) in the variable name.
    Therefore the error is due to the mismatch in variable names.

2.  The corrected code is provided below.

```{r}
library(tidyverse)

ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

3.  Option + Shift + K / Alt + Shift + K pulls up keyboard shortcuts.
    Alternatively, you can get to the same place with Help \> Keyboard Shortcuts Help.

4.  The `my_bar_plot` object is saved because that's what was specified in the `plot` argument of the `ggsave()` call.

```{r}
my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()

my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()

ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
```

```{r}
#| include: false

fs::file_delete("mpg-plot.png")
```
